\section{Conclusion}
\label{sec:conclusion}

\subsection{Key Outcomes}
This project shows that a hybrid architecture, combining deterministic quantitative workflows with LLM-based language generation, can produce a coherent and defensively designed portfolio risk assistant within a prototyping scope. The system computes a broad set of risk metrics from live market data, grounds narrative responses in computed values with a post-generation hallucination check, and retrieves contextual financial knowledge through hybrid lexical-semantic search. The LSTM module contributes forward-looking volatility signals, while intent-conditioned RAG routing improves retrieval precision across query types. Overall, the implementation demonstrates how reliability-critical financial tooling can retain natural language flexibility without surrendering execution control to unconstrained generation.

\subsection{Possible Future Improvements}

\subsubsection{Architecture}
Future work could extend support beyond Yahoo Finance to include additional data sources, broadening the chatbot's applicability to niche equities and cryptocurrency instruments currently excluded from the pipeline. Support for more complex portfolio structures, such as short positions, leveraged products, and options, would also allow the system to serve a wider range of investment strategies.

Replacing session-based storage with a persistent database layer would allow portfolio history, chat records, and analytical outputs to be retained across sessions, enabling longer-term tracking and multi-session continuity.

\subsubsection{LSTM Forecasting Module}
The forecasting module could be improved by incorporating macroeconomic signals such as interest rate data, volatility indices, and yield curve information, since log returns alone are too noisy for robust directional learning. Additional input features would improve the model's ability to anticipate regime shifts and market shocks that are not detectable from the past 60 days of return history alone. Training on real-world portfolio data, rather than synthetically generated sequences, would also improve generalisation across a broader range of portfolio compositions.

\subsubsection{RAG Retrieval Model}
As mentioned in our limitations, expanding the concept knowledge base beyond its current single-chunk-per-concept structure would increase the hard ceiling on Recall@$k$ for concept explanation queries. Adding a cross-encoder re-ranker as a post-retrieval stage would improve chunk ordering before generation, tackling the issue of the current reliance on embedding similarity alone. Fine-tuning the embedding model on financial language would improve retrieval precision for domain-specific terms that general-purpose sentence transformers do not represent well.

\subsubsection{Explanation Layer}
To improve handling of ambiguous multi-intent queries, the current prompt-based intent classifier could be replaced with a fine-tuned BERT model, since the zero-shot LLM classifier occasionally settles on a single route and omits relevant context. Upgrading to a more capable language model better suited to financial reasoning would also be appropriate for any client-facing deployment beyond the testing stage.

\subsubsection{User Interface}
To improve transparency, the interface could surface the system's reasoning steps during loading, showing which tools were called and what context was retrieved before the LLM generated its response. An expandable section in the final output could allow users who want deeper insight to inspect this reasoning trace without cluttering the default view.

\subsection{Closing Remarks}
In summary, this work presents a portfolio risk analyst chatbot that combines a structured quantitative backend, a retrieval-augmented knowledge layer, and an LLM explanation interface. A central design principle is the separation of intent resolution from deterministic execution, which improves reproducibility, observability, and debugging compared with fully autonomous tool planning. The post-generation numeric guard further addresses a critical failure mode in financial LLM responses by checking generated values against computed metrics. The most meaningful next steps are replacing synthetic LSTM training distributions with real portfolio histories, improving retrieval quality with finance-adapted representations and reranking, and adding persistence for cross-session continuity. These extensions would move the system from a strong research prototype toward production-ready decision support.
