\section{Methodology}
\label{sec:methodology}

\subsection{Portfolio construction and notation}
Consider $n$ risky assets. Let $r_t \in \mathbb{R}^n$ denote the vector of simple returns on trading day $t$, and let $w \in \mathbb{R}^n_{\ge 0}$ denote portfolio weights satisfying $\sum_i w_i = 1$ after normalisation. The portfolio return is
\begin{equation*}
r_{p,t} = w^\top r_t .
\end{equation*}
Vectors $w$ are supplied exclusively from the validated interface state and are not inferred by the language model, thereby avoiding implicit reweighting such as undocumented equal-weight assumptions.

\subsection{Risk and performance metrics (current build)}
The implemented pipeline computes a broader diagnostic set, but six core quantities feed the portfolio-level risk scoring layer:
\begin{itemize}
    \item \textbf{Annualised volatility.} The sample standard deviation of $\{r_{p,t}\}$, multiplied by $\sqrt{252}$ under the conventional mapping from daily to annual units for equities.
    \item \textbf{Historical quantile loss.} The empirical 5th percentile of $\{r_{p,t}\}$ at a one-day horizon, reported without scaling to multi-day regulatory value-at-risk horizons; it constitutes a descriptive tail measure rather than a full loss-distribution model.
    \item \textbf{Sharpe ratio.} The ratio of annualised portfolio return in excess of a fixed 4\% risk-free rate to annualised volatility. This is a simplified reference value rather than a user-specific financing-rate adjustment.
    \item \textbf{Maximum drawdown.} Computed on the compounded wealth process $\prod_s (1+r_{p,s})$ as the largest relative decline from a running maximum to a subsequent trough.
    \item \textbf{Average pairwise correlation.} The arithmetic mean of off-diagonal entries in the sample correlation matrix of asset simple returns.
    \item \textbf{Concentration (Herfindahl).} $\sum_i w_i^2$, equal to unity under perfect concentration and to $1/n$ under equal weighting.
\end{itemize}
These six scalars enter a weighted rule-based mapping to a composite score in $[0,1]$ and to ordered risk categories (\texttt{Low}, \texttt{Medium}, \texttt{High}). Thresholds are chosen for demonstration and narrative coherence; they are not calibrated against regulatory templates.

In addition, the quant layer computes supporting diagnostics for explanation and follow-up queries, including conditional value at risk, volatility of volatility, Sortino ratio, beta against \texttt{SPY}, skewness, excess kurtosis, and per-asset risk contribution. These enrich the explanatory layer, but they are distinct from the six inputs used in the risk-scoring rule set.

\subsection{Machine-learning components and limitations}
The current system combines deterministic portfolio analytics with a pre-trained LSTM module for forward-looking signals. In the deployed workflow, the LSTM is invoked for full-analysis and trend-prediction intents to produce a forecast volatility level, a directional prediction (\texttt{Up}/\texttt{Down}), and an associated confidence score. By contrast, the portfolio risk bands (\texttt{Low}/\texttt{Medium}/\texttt{High}) are not produced by a trained classifier; they remain rule-based outputs derived from the six-metric composite score described above.

This separation is intentional. Forecasting is treated as an auxiliary predictive module, whereas the current risk label remains directly auditable from deterministic calculations and explicit thresholds. Any future expansion toward learned risk taxonomies would require a fuller treatment of training data design, temporal leakage control, validation protocol, and calibration.

\subsection{Retrieval-augmented generation}
The RAG subsystem partitions background knowledge into thematic stores (including, inter alia, ticker disambiguation aids, firm-level summaries, macroeconomic indicators, curated definitions, and strategic-allocation notes). Retrieval is conditioned on the inferred dialogue intent: only subsets of corpora are queried for a given turn. Retrieval itself employs a hybrid of lexical matching (Okapi BM25) and dense retrieval of sentence-transformer embeddings persisted in Chroma. Retrieved chunks then pass through relevance filtering and citation attachment before being exposed to the explanation layer. Logging supports retrospective measurement of retrieval quality (e.g.\ recall at fixed $k$ and mean reciprocal rank).

In the \textbf{reference Streamlit deployment} (\texttt{chatbot.py}), deterministic Python assembles the structured context: validated holdings, computed metrics, benchmark labels, risk contributions, optional forecast outputs, recent chat history, and retrieved knowledge-base excerpts. A Gemini explanation model is then called once to generate the final user-facing response from that grounded context. Relative to unconstrained generation, this design narrows-without eliminating-hallucination risk by constraining the model to locally supplied evidence and explicitly authorised numerical values, subject to corpus coverage and chunk boundaries.

\subsection{Agent and intent orchestration}
Each non-empty user utterance receives a primary intent label from Gemini among holistic portfolio analysis, single-metric interrogation, conceptual explanation, trend or outlook questioning, conversational follow-up, and general dialogue. Optional secondary intents and normalised entity fields (metrics, concepts) may accompany the classification.

Dispatch logic implemented in Python branches on the primary label. Computationally intensive paths reuse a single price fetch and return construction step where feasible; explanatory paths may issue retrieval calls with intent tags recognised by the RAG layer (\texttt{full\_analysis}, \texttt{concept\_explanation}, \texttt{trend\_prediction}). Follow-up handling relies on recent session history rather than long-horizon memory. Separation of probabilistic intent classification from deterministic numerical execution is intentional: portfolio arithmetic remains auditable code, while the language model is used primarily for routing and explanation rather than for financial calculation itself.
